### 2014-10-15 doc/thesis/system.tex
Fixing strings in the generated code

\begin{figure}[!ht]
\textbf{Type Language}\\
\dstart
$$
\begin{array}{rlr}
\multicolumn{3}{c}{\textbf{First-level types}}\\
t ::= & \;\; l & \textit{literal types}\\
& | \; b & \textit{base types}\\
& | \; \Nil & \textit{nil type}\\
& | \; \Value & \textit{top type}\\
& | \; \Any & \textit{dynamic type}\\
& | \; \Self & \textit{self type}\\
& | \; t \cup t & \textit{disjoint union types}\\
& | \; p \rightarrow r & \textit{function types}\\
& | \; \{v_{i}, ..., v_{n}\}_{closed|open|unique} & \textit{table types}\\
& | \; x & \textit{type variables}\\
& | \; \mu x.t & \textit{recursive types}\\
& | \; \pi_{i} & \textit{projection types}\\
l ::= & \False \; | \; \True \; | \; {\it int} \; | \; {\it float} \; | \; {\it string} & \\
b ::= & \Boolean \; | \; \Integer \; | \; \Number \; | \; \String & \\
v ::= & k:t \; | \; \Const \; k:t & \textit{field types}\\ 
k ::= & l \; | \; b \; | \; \Any & \textit{key types}\\
\multicolumn{3}{c}{}\\
\multicolumn{3}{c}{\textbf{Second-level types}}\\
s ::= & \;\; t &\\
& | \; t* & \textit{variadic types}\\
& | \; t \times s & \textit{tuple types}\\
p ::= & \Void \; | \; s & \textit{parameters list type}\\
r ::= & \Void \; | \; s \; | \; s \sqcup r & \textit{return types}\\
\end{array}
$$
\dend
\caption{The abstract syntax of Typed Lua types}
\label{fig:typelang}
\end{figure}

@@

\begin{figure}[!ht]
\textbf{Abstract Syntax}\\
\dstart
$$
\begin{array}{rl}
s ::= & \;\; \mathbf{skip} \; | \;
s_{1} \; ; \; s_{2} \; | \;
\vec{l} = \vec{e}  \; | \;
\mathbf{while} \; e \; \mathbf{do} \; s \; | \;
\mathbf{if} \; e \; \mathbf{then} \; s_{1} \; \mathbf{else} \; s_{2}\\
& | \; \mathbf{local} \; \vec{n{:}t} = \vec{e} \; \mathbf{in} \; s \; | \;
\mathbf{rec} \; n{:}t = f \; \mathbf{in} \; s \; | \;
\mathbf{return} \; \vec{e} \; | \;
a\\
e ::= & \;\; \mathbf{nil} \; | \;
\mathbf{false} \; | \;
\mathbf{true} \; | \;
{\it number} \; | \;
{\it string} \; | \;
{...} \; | \;
f \; | \;
\{ \; \vec{c} \; \} \\
& | \; e_{1} + e_{2} \; | \;
e_{1} \; {..} \; e_{2} \; | \;
e_{1} == e_{2} \; | \;
e_{1} < e_{2} \; | \;
e_{1} \; \mathbf{and} \; e_{2} \; | \;
e_{1} \; \mathbf{or} \; e_{2} \\
& | \; \mathbf{not} \; e \; | \;
- e \; | \;
\# \; e \; | \;
a \; | \;
l \; | \;
(e) \; | \;
{<}t{>} \; e\\
l ::= & \; n \; | \;
e_{1}[e_{2}]\\
a ::= & \; e_{1}(\vec{e_{2}}) \; | \;
e_{1}{:}n(\vec{e_{2}})\\
f ::= & \; \mathbf{fun} \; (){:}r \; s \; | \;
\mathbf{fun} \; ({...}{:}t){:}r \; s \; | \;
\mathbf{fun} \; (\vec{n{:}t}){:}r \; s \; | \;
\mathbf{fun} \; (\vec{n{:}t},{...}{:}t){:}r \; s\\
c ::= & \; [e_{1}] = e_{2}\\
n ::= & \; {\it name}\\
\end{array}
$$
\dend
\caption{Typed Lua abstract syntax}
\label{fig:syntax}
\end{figure}

>>>>

\begin{figure}[!ht]
\textbf{Type Language}\\
\dstart
$$
\begin{array}{rlr}
\multicolumn{3}{c}{\textbf{First-level types}}\\
t ::= & \;\; l & \textit{literal types}\\
& | \; b & \textit{base types}\\
& | \; \Nil & \textit{nil type}\\
& | \; \Value & \textit{top type}\\
& | \; \Any & \textit{dynamic type}\\
& | \; \Self & \textit{self type}\\
& | \; t \cup t & \textit{disjoint union types}\\
& | \; p \rightarrow r & \textit{function types}\\
& | \; \{v_{i}, ..., v_{n}\}_{closed|open|unique} & \textit{table types}\\
& | \; x & \textit{type variables}\\
& | \; \mu x.t & \textit{recursive types}\\
& | \; \pi_{i} & \textit{projection types}\\
l ::= & \;\; \False \; | \; \True \; | \; {\it int} \; | \; {\it float} \; | \; {\it string} & \\
b ::= & \;\; \Boolean \; | \; \Integer \; | \; \Number \; | \; \String & \\
v ::= & \;\; k:t \; | \; \Const \; k:t & \textit{field types}\\ 
k ::= & \;\; l \; | \; b \; | \; \Any & \textit{key types}\\
\multicolumn{3}{c}{}\\
\multicolumn{3}{c}{\textbf{Second-level types}}\\
s ::= & \;\; t &\\
& | \; t* & \textit{variadic types}\\
& | \; t \times s & \textit{tuple types}\\
p ::= & \;\; \Void \; | \; s & \textit{parameters list type}\\
r ::= & \;\; \Void \; | \; s \; | \; s \sqcup r & \textit{return types}\\
\end{array}
$$
\dend
\caption{The abstract syntax of Typed Lua types}
\label{fig:typelang}
\end{figure}

@@

In this section we will use a reduced core of Typed Lua to present
its typing rules in an easier way.
It is a reduced core of the Lua language with explicit type annotations,
without syntactic sugar, with only if and while control flow statements,
and making variable scope explicit.

\begin{figure}[!ht]
\textbf{Abstract Syntax}\\
\dstart
$$
\begin{array}{rl}
s ::= & \;\; \mathbf{skip} \; | \;
s_{1} \; ; \; s_{2} \; | \;
\vec{l} = el  \; | \;
\mathbf{while} \; e \; \mathbf{do} \; s \; | \;
\mathbf{if} \; e \; \mathbf{then} \; s_{1} \; \mathbf{else} \; s_{2}\\
& | \; \mathbf{local} \; \vec{n{:}t} = el \; \mathbf{in} \; s \; | \;
\mathbf{rec} \; n{:}t = f \; \mathbf{in} \; s \; | \;
\mathbf{return} \; el \; | \;
a\\
e ::= & \;\; \mathbf{nil} \; | \;
\mathbf{false} \; | \;
\mathbf{true} \; | \;
{\it int} \; | \;
{\it float} \; | \;
{\it string} \; | \;
{...} \; | \;
f \; | \;
\{ \; cl \; \} \\
& | \; e_{1} + e_{2} \; | \;
e_{1} \; {..} \; e_{2} \; | \;
e_{1} == e_{2} \; | \;
e_{1} < e_{2} \; | \;
e_{1} \; \mathbf{and} \; e_{2} \; | \;
e_{1} \; \mathbf{or} \; e_{2} \\
& | \; \mathbf{not} \; e \; | \;
- e \; | \;
\# \; e \; | \;
a \; | \;
l \; | \;
{<}t{>} \; n\\
l ::= & \;\; n \; | \;
e_{1}[e_{2}]\\
el ::= & \;\; \mathbf{nothing} \; | \;
\vec{e} \; | \;
me \; | \;
\vec{e}, me\\
cl ::= & \;\; \mathbf{nothing} \; | \;
\vec{c} \; | \;
\vec{e} \; | \;
me \; | \;
\vec{c}, \vec{e} \; | \;
\vec{c}, me \; | \;
\vec{c}, \vec{e}, me \; | \;
\vec{e}, me\\
me ::= & \;\; a \; | \;
{...}\\
a ::= & \;\; e_{1}(\vec{e_{2}}) \; | \;
e_{1}{:}n(\vec{e_{2}})\\
f ::= & \;\; \mathbf{fun} \; (){:}r \; s \; ; \; \mathbf{return} \; el \; | \;
\mathbf{fun} \; ({...}{:}t){:}r \; s \; ; \; \mathbf{return} \; el\\
& | \; \mathbf{fun} \; (\vec{n{:}t}){:}r \; s \; ; \; \mathbf{return} \; el \; | \;
\mathbf{fun} \; (\vec{n{:}t},{...}{:}t){:}r \; s \; ; \; \mathbf{return} \; el\\
c ::= & \;\; [e_{1}] = e_{2}\\
n ::= & \;\; {\it name}\\
\end{array}
$$
\dend
\caption{Typed Lua abstract syntax}
\label{fig:syntax}
\end{figure}

Figure \ref{fig:syntax} presents the abstract syntax of core Typed Lua.
It splits the syntactic categories as follows:
$s$ are statements, $e$ are expressions, $l$ are left-hand values,
$el$ are expression lists, $c$ are constructor fields, $cl$ are lists of
constructor fields, $me$ are multiple expressions, $a$ are function
applications, $f$ are function declarations, $n$ are variable names,
$t$ are first-level types, and $r$ are second-level types.
The notation $\vec{n{:}t}$ denotes the non-empty list
$n_{1}{:}t_{1}, ..., n_{n}{:}t_{n}$.

We will present the typing rules as a deduction system for the typing
relations $\env_{1} \vdash s : \env_{2}$ and $\env_{1} \vdash e : t, \env_{2}$. 
We will use the first one for typing statements, and the sencond one
for typing expressions. 
The first relation means that given the type environment $\env_{1}$,
typing the statement $s$ produces the new type environment $\env_{2}$. 
The sencond relation means that given the type environment $\env_{1}$,
the expression $e$ has type $t$, and produces the new type environment $\env_{2}$.
We need two typing relations because only expressions have types,
though both expressions and statements can modify the environment.
The environment maps variables to types, and we will use $\env_{1}[n \mapsto t]$
for meaning that we extended the environment $\env_{1}$ with the
variable $n$ that maps to type $t$, and this mapping does not affect
the other mappings that extisted before.
